---
title: "Data_606_Lab_6_Inference for categorical data.rmd"
author: "Enid Roman"
date: "2022-10-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(500)
library(tidyverse)
library(openintro)
library(infer)
```
```{r}
data('yrbss', package='openintro')
print(head(yrbss,2))
```
### Exercise 1 

#### The counts within each category for the amount of days these students have texted while driving within the past 30 days are:
#### 4792 have reported 0 days.
#### 4646 have reported did not drive. 
#### 925 have reported drive 1 to 2 days.
#### 918 have reported NA days
#### 827 have reported 30 days. 
#### 493 have reported 3 to 5 days.
#### 373 have reported 10 to 19 days.
#### 311 have reported 6 to 9 days. 
#### 298 have reported 20 to 29 days. 

```{r}
yrbss %>%
  count(text_while_driving_30d, sort=TRUE)
  
```

### Exercise 2

#### 6.64% (463/6977) (if counting the NA) is the proportion of people who have texted while driving every day in the past 30 days and never wear helmets. (See below)

```{r}
data('yrbss', package='openintro')
no_helmet <- yrbss %>%
  filter(helmet_12m == "never") 
no_helmet <- no_helmet %>%
  mutate(text_ind = ifelse(text_while_driving_30d == "30", "yes", "no"))
no_helmet %>%
  count(text_ind)
```
```{r}
no_helmet %>%
filter(!is.na(text_ind)) %>%
filter(helmet_12m == "never") %>%
filter(text_ind == "yes") %>%
nrow() / nrow(no_helmet)
```
```{r}
no_helmet %>%
filter(text_ind != "") %>%
specify(response = text_ind, success = "yes") %>%
generate(reps = 1000, type = "bootstrap") %>%
calculate(stat = "prop") %>%
get_ci(level = 0.95)
```
### Exercise 3

#### We are 95% confident that the proportion of non-helmet wearers that have texted while driving each day for the past 30 days based on this survey is between 6.57% and 7.75%.

#### The margin of error for the estimate of the proportion of non-helmet wearers that have texted while driving each day for the past 30 days based on this survey is .0063 or 6%


```{r}
1.96 * sqrt((0.0775*(1-.0775)/6977))
```
### Exercise 4 

#### Hours TV Per School Day

#### Proportion of Interest of students who reported watching less than 1 hr of tv per school day is 15.96% (2168/13583).
#### We are 95% confident that the proportion of students who reported watching less than 1 hr of tv per school day is between 15.72% and 17.04%.
#### The margin of error for the estimate of the proportion of students who reported watching less than 1 hr of tv per school day is .0064 or 6%.

#### School Night Hours of Sleep

#### Proportion of Interest of students who reported less than 5hrs of sleep on school nights is 7.15% (965/12335).
#### We are 95% confident that the proportion of students who reported less than 5hrs of sleep on school nights is between 7.36% and 8.30%. 
#### The margin of error for the estimate of the proportion of students who reported less than 5hrs of sleep on school nights is .0048 or 5%.


```{r}
yrbss %>%
  count(hours_tv_per_school_day, sort=TRUE)
```

```{r}
tv<- yrbss %>%
  filter(!is.na(hours_tv_per_school_day)) %>%
  mutate(tv_ind = ifelse(hours_tv_per_school_day == "<1", "yes", "no"))

tv %>%
  count(tv_ind)
```


```{r}

tv %>%
    specify(response = tv_ind, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

```
```{r}
1.96 * sqrt((.1704*(1-.1704)/13245))
```
#### School Night Hours Sleep

```{r}
yrbss %>%
  count(school_night_hours_sleep, sort=TRUE)
```

```{r}
sleep <- yrbss %>%
  filter(!is.na(school_night_hours_sleep)) %>%
  mutate(sleep_ind = ifelse(school_night_hours_sleep == "<5", "yes", "no"))

sleep %>%
  count(sleep_ind)
```

```{r}
sleep %>%
 specify(response = sleep_ind, success = "yes") %>%
 generate(reps = 1000, type = "bootstrap") %>%
 calculate(stat = "prop") %>%
 get_ci(level = 0.95)
```
```{r}
1.96 * sqrt((.0831*(1-.0831)/12335))
```

```{r}
n <- 1000
```

```{r}
p <- seq(from = 0, to = 1, by = 0.01)
me <- 2 * sqrt(p * (1 - p)/n)
p
```

```{r}
dd <- data.frame(p = p, me = me)
ggplot(data = dd, aes(x = p, y = me)) + 
  geom_line() +
  labs(x = "Population Proportion", y = "Margin of Error")
```

### Exercise 5 

#### The relationship between p and me is the margin of error increases as the population proportion increases. Margin of error is greatest at the population of 50%. For a given sample size for which a value of p is margin of error maximized is at .5.

### Exercise 6 (Not sure which app you were referring to)

#### The distribution of sampling proportions with sampling size of 300 and p=0.1, center is at .01 and spread conformsis .08 to .11.

```{r}
p <- 0.1
n <- 300

(p*(1-p)/n)^.5
```
```{r}
.1-(p*(1-p)/n)^.5
```
```{r}
.1+(p*(1-p)/n)^.5
```
### Exercise 7

#### Keeping the n constant and changing p the shape and center does change. The spread of the sampling distribution does vary as p changes. The distribution of sampling proportions with sampling size of 300 and p=0.5, center is at .03 and spread conformsis .07 to .12. Increasing the p does increase the spread. 

```{r}
p <- 0.5
n <- 300

(p*(1-p)/n)^.5
```

```{r}
.1-(p*(1-p)/n)^.5

```

```{r}
.1+(p*(1-p)/n)^.5
```
### Exercise 8

#### Keeping the n constant and changing p the shape and center does not change much. The spread of the sampling distribution does vary as p changes. The distribution of sampling proportions with sampling size of 300 and p=0.5, center is at .03 and spread conformsis .08 to .13. Increasing the n does increase the spread a little.

```{r}
p <- 0.5
n <- 400

(p*(1-p)/n)^.5
```

```{r}
.1-(p*(1-p)/n)^.5
```

```{r}
.1+(p*(1-p)/n)^.5
```


### Exercise 9

#### Yes, there is convincing evidence that those who sleep 10+ hours per day are more likely to strength train every day of the week. The confidence interval range for sleep less than 10 is from 26.04% to 27.67%. The confidence interval range for sleep more than 10 is from 31.63% to 41.86%. Is a aprox. 5% to 15% difference.

```{r}
sleep_less_than_10 <- yrbss %>%
  filter(school_night_hours_sleep != "10+") 

sleep_less_than_10 %>%
  mutate(physical = ifelse(physically_active_7d == 7, "yes", "no")) %>%
  drop_na(physical) %>%
  specify(response = physical, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

```{r}
sleep_10plus <- yrbss %>%
  filter(school_night_hours_sleep == "10+") 

sleep_10plus %>%
  mutate(physical = ifelse(physically_active_7d == 7, "yes", "no")) %>%
  drop_na(physical) %>%
  specify(response = physical, success = "yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```
### Exercise 10

#### The probablity that you could detect a change (at a significance level of 0.05) simply by chance would be 62%. 

```{r}
mu <- 2.5
(cv <- qnorm(0.05, 
             mean=0, 
             sd=1, 
            lower.tail=FALSE))
```


```{r}
pnorm(mu, mean=0, sd=1, lower.tail=FALSE)
```
### Exercise 11

#### With and estimate margin of error no greater than 1% with 95% confidence. I would have to sample 9694 people to ensure that you I am within the guidelines.

#### ME <-  1.96 * SE for 95% confidence margin
#### SE <-  sqrt(p*(1 - p)/n)
#### ME <-  1.96 * sqrt(p*(1 - p)/n)
#### ME^2 <-  1.96^2 * p*(1 - p)/n
#### n <-  1.96^2 * p*(1 - p)/ME^2

```{r}

ME <- 0.01
p <- 0.5

1.96^2 * p *(1 - 0.5)/ME^2
```




