---
title: "Lab 1: Intro to R"
author: "Enid Roman"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tinytex)
library(tidyverse)
library(openintro)
```

###**Arbuthnot**

```{r}
data('arbuthnot', package='openintro')
arbuthnot
```

```{r}
glimpse(arbuthnot)
```

```{r}
arbuthnot$boys
```

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```

```{r}
ggplot(data = arbuthnot, aes(x = year, y = girls)) + 
  geom_point()
```

### Exercise 2

There is an apparent flucturating trend in the number of girls baptized over the years. Between 1650 and 1658 the number of girls baptised dropped tremendously below 3000. Around 1659 to after 1700 the number of girls baptised began to increase again reaching close to 8000. You can see a sudden dip in 1670 and 1740 which picked up quickly.  

```{r trend-girls}
# Insert code for Exercise 2 here

ggplot(data= arbuthnot, aes(x = year, y = girls)) +
 geom_line()
```

```{r}
5218 + 4683
```

```{r}
arbuthnot$boys + arbuthnot$girls
```

```{r}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot
```

```{r}
ggplot(data = arbuthnot, aes(x = year, y = total)) + 
  geom_line()
```

```{r}
5218 / 4683
```

```{r}
arbuthnot <- arbuthnot %>%
  mutate(boy_to_girl_ratio = boys / girls)
arbuthnot
```

```{r}
5218 / (5218 + 4683)
```

```{r}
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)
arbuthnot
```


### Exercise 3

I see extreme flucturation over the years of boys born over time. Each year there is a sudden decrease and a sudden increase over the years. 

```{r plot-prop-boys-arbuthnot}
# Insert code for Exercise 3 here
ggplot(data = arbuthnot, aes(x = year, y = boy_ratio)) + 
  geom_line()
```

```{r}
arbuthnot <- arbuthnot %>%
  mutate(more_boys = boys > girls)
arbuthnot
```

```{r}
arbuthnot %>%
  summarize(min = min(boys), max = max(boys))
```

###**Present**

```{r}
data('present', package='openintro')
present
```

### Exercise 4

The years that are included is 1940 to 2002. The dimensions of the data frame is 63 rows and 3 columns. The variable (column) names are year, boys, and girls. 

```{r dim-present}
# Insert code for Exercise 4 here
range(present$year)
```
Continue for Exercise 4

```{r}
glimpse(present)
```

### Exercise 5

No, the the counts compare to Arbuthnot's count are not of a similar magnitute. The total counts of boys and girls for the present dataset is over 4,000,000 compare to the total counts of boys and girls for the Arbuthnot's dataset. 

```{r count-compare}
# Insert code for Exercise 5 here
present$boys + present$girls
```

Continue for Exercise 5

```{r}
present <- present %>%
  mutate(total = boys + girls)
present
```
### Exercise 6

I see a decrease in ratio of boys to girls born over time. Even though there is a decrease in ratio of boy to girl born over time, Arbuthnot's observation about boys being born in greater poportion than girls hold up in the U.S.

```{r}
present <- present %>%
  mutate(boy_to_girl_ratio = boys / girls)
present
```
Continue for Exercise 6

```{r}
present <- present %>%
  mutate(boy_ratio = boys / total)
present
```
Continue for Exercise 6

```{r}
present <- present %>%
  mutate(more_boys = boys > girls)
present
```
Continue for Exercise 6

```{r}
ggplot(data = present, aes(x = year, y = boy_ratio)) + 
  geom_line()
```

### Exercise 7


The year that we see with the most total number of births in the U.S. is 1961.

```{r find-max-total}
# Insert code for Exercise 7 here
present <- present %>%
  mutate(present$boys + present$girls)
present <- present %>%
  arrange(desc(total))
head(present)

```

